---
title: 'Data Analysis Assignment #1 (50 points total)'
author: "Mark, Ryan"
output:
  html_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```
### Project Overview: ###

***This project will examine sample data collected from a large marine sea snail called Abalone. The information that we have retrieved is based on an observational study that was performed on Abalones. The objective of that study was to be able to predict the age of Abalones based on their physical measurements. At the conclusion of that study it was determined that it was unsuccessful at meeting its overall objective. The intent of this exploratory analysis is to determine plausible reasons why the original study was unsuccessful in predicting abalone age based on physical characteristics.***

### Observations: ###

***At the conclusion of the data analysis it is observed that the sample data possess outliers across several comparisons. A box plot of sex compared to ratios indicated that each sex had outliers present. Additionally, these outliers appeared to skew the histogram of sex by ratio. This presented the overall data with a right-skew which indicates non-normality. When the volume and whole was measured against the age class in section 4, we see a similar occurrence, outliers are present. When volume and whole are compared to rings, we see a great deal of dispersion. Particularly in the upper ranges of ring counts where there is a noticeable amount of variation. In section 5, the linear relationships were seen across all sexes. For example, it was observed that when measuring volume by age class that as the volume increased so did the age class. This indicates (based on the sample data collected) that the volume size of the Abalones collected could be a predictor of age. Further statistical analysis is needed to validate this hypothesis and will likely be addressed in project #2. In summary, it is believed that the core of issues causing the failure to predict age can be attributed to outliers. It is recommended that the outliers be removed for further analysis so that we can observe the data within the normal parameters.***

```{r analysis_setup1, message = FALSE, warning = FALSE}

# a) Load the ggplot2 and gridExtra packages.

library(ggplot2)
#library(gridExtra)
library(knitr)

# b) Use read.csv() to read the abalones.csv into R, assigning the data frame to "mydata."

mydata <- read.csv("abalones.csv", sep = ",")


# c) Use the str() function to verify the structure of "mydata." You should have 1036 observations
# of eight variables.

str(mydata)

# d) Define two new variables, VOLUME and RATIO. Use the following statements to define VOLUME and
# RATIO as variables appended to the data frame "mydata."

mydata$VOLUME <- mydata$LENGTH * mydata$DIAM * mydata$HEIGHT
mydata$RATIO <- mydata$SHUCK / mydata$VOLUME

```
### Test Items starts from here - There are 6 sections ##########################

##### Section 1: (6 points) Summarizing the data.

(1)(a) (1 point) Use *summary()* to obtain and present descriptive statistics from mydata.  Use table() to present a frequency table using CLASS and RINGS. There should be 115 cells in the table you present.  

### Summary of Data Findings: ###

***The summary data() indicates that we have 10 categories of data. Additionally, there are classification level types of  data. For example, Sex is listed as a classification type data and we can observe that the sample contains 326 Females, 329 Infants, and 381 Males. There are 5-types of Classes that are within the data listed from A1 – A5. The remaining 8 categories contain measurement (numerical data) where minimum, maximum, median, and quartile values can be observed.***
```{r Part_1a}
summary(mydata)
table(mydata$CLASS,mydata$RINGS)
```
**Question (1 point):  Briefly discuss the variable types and distributional implications such as potential skewness and outliers.**

***Answer: Once the summary function is used, I am able to identify that “Class” has 5 variables (A1, A2, A3, A4, A5). The table function allows me to see the distribution by Class variable. For example, I am able to see that A1 appears 9-times for rings that are equal to 3. Also, based on the frequency I am able to see that Class A3 has the largest frequency at 9-rings (peak of distribution). Also, the descriptive statistics for Rings indicate that the median is = 9 and the mean is equal to 9.993. This indicates that the distribution for Rings has a positive skew or is right-skewed. Also, I am able to identify that the max frequency for Rings is 25 and that there is one occurrence for this = Class A5.The upper range for ring counts reflect lower frequency counts which may potentially indicate outliers.***

(1)(b) (1 point) Generate a table of counts using SEX and CLASS. Add margins to this table (Hint: There should be 15 cells in this table plus the marginal totals. Apply *table()* first, then pass the table object to *addmargins()* (Kabacoff Section 7.2 pages 144-147)).  Lastly, present a barplot of these data; ignoring the marginal totals.  

```{r Part_1b}
par(bg="gray98")
a<- table(mydata$SEX,mydata$CLASS)
addmargins(a)
barplot(a,col=c("yellow","blue","green"),horiz=TRUE,main = "Count of Abalone sex by class",legend=TRUE)
```

**Essay Question (2 points):  Discuss the sex distribution of abalones.  What stands out about the distribution of abalones by CLASS?**

***Answer: There are (3) types of abalone sexes across (5) types of Classes. By using “add margins” I am able to identify that the sex contains F (326), I (329) and M (381). Therefore, the largest group under sex is M (381). The distribution of abalones by class indicates that the largest class for these sexes is A3 (total of sexes), with A1 being the smallest. The highest count for F by Class occurs in A3. The highest count for I by Class occurs in A2. The highest count for M by Class occurs in A3. We can conclude that Class A3 contains the highest count for Sex F and Sex M.***

(1)(c) (1 point) Select a simple random sample of 200 observations from "mydata" and identify this sample as "work." Use *set.seed(123)* prior to drawing this sample. Do not change the number 123. Note that *sample()* "takes a sample of the specified size from the elements of x." We cannot sample directly from "mydata." Instead, we need to sample from the integers, 1 to 1036, representing the rows of "mydata." Then, select those rows from the data frame (Kabacoff Section 4.10.5 page 87). 

Using "work", construct a scatterplot matrix of variables 2-6 with *plot(work[, 2:6])* (these are the continuous variables excluding VOLUME and RATIO). The sample "work" will not be used in the remainder of the assignment. 

```{r Part_1c}
set.seed(123)
work <- mydata[sample(1:nrow(mydata),200, replace=FALSE),]
plot(work[,2:6],col=c("yellow","red","blue","purple","green"),main="Simple Random Sample of 200 Obervations")
```

-----

##### Section 2: (5 points) Summarizing the data using graphics.

(2)(a) (1 point) Use "mydata" to plot WHOLE versus VOLUME. Color code data points by CLASS.

### Scatterplot Observations: ###

***The Abalone scatterplot for volume by whole represents a positive slope (linear relationship). When we examine the individual classes within the scatterplot it is noted that outliers can be also observed. Particularly, this can be seen with class A5 and the data points that are the upper right-corner of the scatterplot. The shuck vs whole scatterplot also represents a positive linear relationship. In this graph we have it has been observed that the lower value in shuck and whole reflect a tighter positive relationship, however, as that count increases (both x & y) we observe that the shuck and whole have a larger cone (slightly less linear) relationship.***

```{r Part_2a}
par(bg="gray98")
plot(mydata$WHOLE,mydata$VOLUME,col=mydata$CLASS, pch=c(16,19,21),xlab="Whole",ylab="Volume",main="Abalone Scatter Plot (Volume by Whole)")
legend(x=0,y=1000,legend=c("A1","A2","A3","A4","A5"),fill=c(1,2,3,4,5))
```


(2)(b) (2 points) Use "mydata" to plot SHUCK versus WHOLE with WHOLE on the horizontal axis. Color code data points by CLASS. As an aid to interpretation, determine the maximum value of the ratio of SHUCK to WHOLE.  Add to the chart a straight line with zero intercept using this maximum value as the slope of the line. If you are using the 'base R' *plot()* function, you may use *abline()* to add this line to the plot. Use *help(abline)* in R to determine the coding for the slope and intercept arguments in the functions. If you are using ggplot2 for visualizations, *geom_abline()* should be used.

```{r Part_2b}
par(bg="gray98")
plot(mydata$WHOLE,mydata$SHUCK,col=mydata$CLASS, pch=c(16,19,21),,xlab="Whole",ylab="Shuck",main="Abalone Scatter Plot (Shuck by Whole)")
max(mydata$SHUCK/mydata$WHOLE)
abline(a= 0,b= .562,col="red")
legend(x=0,y=160,legend=c("A1","A2","A3","A4","A5"),fill=c(1,2,3,4,5))
```

**Essay Question (2 points):  How does the variability in this plot differ from the plot in (a)?  Compare the two displays.  Keep in mind that SHUCK is a part of WHOLE.  Consider the location of the different age classes.**

***Answer: Graph “A” displays a wider cone across all data variables and a higher range of values on the y-axis. In graph “B” we see an overall tighter cone across all data variables and a lower range of values on the y-axis. When examining the specific patterns for the class it can be seen that in graph “A” there is larger variation as the data variables increase. As both x and y increase the formation expands and there is a noticeable amount of variation between data points. Similarly, the same observation is seen for graph “B”. In particular, the class assigned to the color “aqua” in graph “B” experiences greater variation near the max ranges of the x and y axis’.***

-----

##### Section 3: (8 points) Getting insights about the data using graphs.

(3)(a) (2 points) Use "mydata" to create a multi-figured plot with histograms, boxplots and Q-Q plots of RATIO differentiated by sex. This can be done using *par(mfrow = c(3,3))* and base R or *grid.arrange()* and ggplot2. The first row would show the histograms, the second row the boxplots and the third row the Q-Q plots. Be sure these displays are legible.  

```{r Part_3a}
par(bg="gray98")
par(mfrow = c(3,3))
hist(mydata$RATIO[mydata$SEX=='M'],main='HIST: Ratio for M',xlab='Ratio for M',col=c("red","yellow","green"))
hist(mydata$RATIO[mydata$SEX=='F'],main='HIST: Ratio for F',xlab='Ratio for F',col=c("lightblue","orange","purple"))
hist(mydata$RATIO[mydata$SEX=='I'],main='HIST: Ratio for I',xlab='Ratio for I',col=c("aquamarine","navyblue","violet"))
boxplot(mydata$RATIO[mydata$SEX=='M'],main='Box Plot: Ratio for M',xlab='Ratio for M',col=c("yellow"))
boxplot(mydata$RATIO[mydata$SEX=='F'],main='Box Plot: Ratio for F',xlab='Ratio for F',col=c("green"))
boxplot(mydata$RATIO[mydata$SEX=='I'],main='Box Plot: Ratio for I',xlab='Ratio for I',col=c("orange"))
qqnorm(mydata$RATIO[mydata$SEX=='M'],main='Q-Q Plot: Ratio for M',xlab='Ratio for M',col=c("darkgreen"))
qqnorm(mydata$RATIO[mydata$SEX=='F'],main='Q-Q Plot: Ratio for F',xlab='Ratio for F',col=c("purple"))
qqnorm(mydata$RATIO[mydata$SEX=='I'],main='Q-Q Plot: Ratio for I',xlab='Ratio for I',col=c("blue"))
```

**Essay Question (2 points): Compare the displays.  How do the distributions compare to normality? Take into account the criteria discussed in the sync sessions to evaluate non-normality.**

***Answer: Comparing the histograms in row 1 I observed that the shape is not symmetric. I can tell that all 3 histograms are right skewed indicated by the relationship of the peak values to the tail. Also, when we look for normality, we can also view the box plots in row 2 for reference. Here we see that each graph has outliers. Additionally, if I wanted review this process in detail, I could also run a Skewness and Kurtosis test.***

(3)(b) (2 points) Use the boxplots to identify RATIO outliers (mild and extreme both) for each sex. Present the abalones with these outlying RATIO values along with their associated variables in "mydata" (Hint:  display the observations by passing a data frame to the kable() function).  

```{r Part_3b}
par(bg="gray98")
par(mfrow = c(1,3))
boxplot(mydata$RATIO[mydata$SEX=='M'],main='Box Plot: Ratio for M',xlab='Ratio for M',col=c("yellow"))
boxplot(mydata$RATIO[mydata$SEX=='F'],main='Box Plot: Ratio for F',xlab='Ratio for F',col=c("green"))
boxplot(mydata$RATIO[mydata$SEX=='I'],main='Box Plot: Ratio for I',xlab='Ratio for I',col=c("orange"))

M <- boxplot.stats(mydata$RATIO[mydata$SEX=='M'])[["out"]]
F <- boxplot.stats(mydata$RATIO[mydata$SEX=='F'])[["out"]]
I <- boxplot.stats(mydata$RATIO[mydata$SEX=='I'])[["out"]]

t <- match(c(M,F,I),mydata$RATIO)
p <- mydata[t,]
kable(p)
#summary(p)
```

**Essay Question (2 points):  What are your observations regarding the results in (3)(b)?**

***Answer: 19 outliers are captured in the results. 6 results for Females, 8 for Infants, and 5 for Males. In the classes I observed the following outliers, A1 = 6, A2=3, A3=9, A4=0, A5=1. The max ratio outlier is .31176 which occurred in class A2 (Female) and the min ratio outlier is .06734 which occurred in class A5 (Female).***

-----

##### Section 4: (8 points) Getting insights about possible predictors.

(4)(a) (3 points) With "mydata," display side-by-side boxplots for VOLUME and WHOLE, each differentiated by CLASS There should be five boxes for VOLUME and five for WHOLE. Also, display side-by-side scatterplots:  VOLUME and WHOLE versus RINGS. Present these four figures in one graphic:  the boxplots in one row and the scatterplots in a second row. Base R or ggplot2 may be used.


```{r Part_4a}
par(bg="gray98")
par(mfrow = c(2,2))
boxplot(mydata$VOLUME~mydata$CLASS,ylab='VOLUME',col=c("yellow","red","green","orange","purple"),main= "Volume by Class")
boxplot(mydata$WHOLE~mydata$CLASS,ylab='WHOLE',col=c("darkgreen","pink","aquamarine","blue","salmon"),main="Whole by Class")
plot(mydata$VOLUME~mydata$RINGS,ylab='VOLUME',xlab='RINGS',col=c("darkgreen","pink","aquamarine","blue","salmon"),main="Volume by Rings")
plot(mydata$WHOLE~mydata$RINGS,ylab='WHOLE',xlab='RINGS',col=c("yellow","red","green","orange","purple"),main="Whole by Rings")
```

**Essay Question (5 points) How well do you think these variables would perform as predictors of age?  Explain.**

***Answer: To better understand if these variables can be a predictor of age, we need to examine the correlation. For example, volume encompasses the product of length, diameter, and height and is represented on the y-axis (Graph in row 1, col 1). Then on the x-axis there are the classes (A1 – A5). In this particular instance we can observe that the median increases as the class number increases. This would indicate that the median values increase across the range of classes. Therefore, if age was correlated with volume according to size then this may offer insight into the age of an abalone based on the hypothesis that greater volume correlates to age (positive relationship). Likewise, when we use the same comparison of volume and rings, we see a positive relationship. However, it should be noted as the x and y values increase on the volume to rings and whole to rings, we see variability in the data points. Greater dispersion is shown which indicates less of a correlation or accuracy. This would indicate that a predictor of age may be more feasible when the rings are less than 15. Once the rings are greater than 15 there is more variability making it less accurate to predict the age.***

-----

##### Section 5: (12 points) Getting insights regarding different groups in the data.

(5)(a) (2 points) Use *aggregate()* with "mydata" to compute the mean values of VOLUME, SHUCK and RATIO for each combination of SEX and CLASS. Then, using *matrix()*, create matrices of the mean values. Using the "dimnames" argument within *matrix()* or the *rownames()* and *colnames()* functions on the matrices, label the rows by SEX and columns by CLASS. Present the three matrices (Kabacoff Section 5.6.2, p. 110-111). The *kable()* function is useful for this purpose.  You do not need to be concerned with the number of digits presented.

```{r Part_5a}
a <- aggregate(mydata$VOLUME,by=list(mydata$SEX, mydata$CLASS),FUN=mean)
b <- aggregate(mydata$SHUCK,by=list(mydata$SEX, mydata$CLASS),FUN=mean)
c <- aggregate(mydata$RATIO,by=list(mydata$SEX, mydata$CLASS),FUN=mean)

am <- matrix(round(a$x,2),nrow=3,ncol=5,dimnames = list(c("Female","Infant","Male"),
                                       c("A1","A2","A3","A4","A5")))
bm <- matrix(round(b$x,2),nrow=3,ncol=5,dimnames = list(c("Female","Infant","Male"),
                                       c("A1","A2","A3","A4","A5")))
cm <- matrix(round(c$x,4),nrow=3,ncol=5,dimnames = list(c("Female","Infant","Male"),
                                       c("A1","A2","A3","A4","A5")))

library(knitr)
kable(list(am,bm,cm))
```

(5)(b) (3 points) Present three graphs. Each graph should include three lines, one for each sex. The first should show mean RATIO versus CLASS; the second, mean VOLUME versus CLASS; the third, mean SHUCK versus CLASS. This may be done with the 'base R' *interaction.plot()* function or with ggplot2 using *grid.arrange()*.

```{r Part_5b, fig.width = 9}
par(bg="gray98")
interaction.plot(x.factor = mydata$CLASS, trace.factor = mydata$SEX,response = mydata$RATIO,
                 col=c("blue","red","black"),lty=1,lwd=2,xlab="Class",ylab="Ratio",trace.label = "SEX",main="Abalone Ratio by Class")

interaction.plot(x.factor = mydata$CLASS, trace.factor = mydata$SEX,response = mydata$VOLUME,
                 col=c("blue","red","black"),lty=1,lwd=2,xlab="Class",ylab="Volume",trace.label = "SEX",main="Abalone Volume by Class")

interaction.plot(x.factor = mydata$CLASS, trace.factor = mydata$SEX,response = mydata$SHUCK,
                 col=c("blue","red","black"),lty=1,lwd=2,xlab="Class",ylab="Shuck",trace.label = "SEX",main="Abalone Shuck by Class")


```

**Essay Question (2 points):  What questions do these plots raise?  Consider aging and sex differences.**

***Answer: The Ratio to Class line graph indicates a negative relationship. That is, for all sexes, as they decrease the class size increases. Class A1 yields the highest ratio while Class A5 yields the lowest ratios. Questions that arise are: 1) Why are Infants ratio to class so linear compared to Male and Female sexes? The Ratio to Class line graph indicates a positive relationship. All sexes increase in volume as Class increases. Questions that arise: 1) Why are Female abalones larger compared to Male and Infants? 2) What factors are causing all three abalones to peak in class A4 but then maintain a steady volume through class A5?Similar to the Ratio and Class line graph, the Shuck and Class indicates a positive relationship. All sexes increase in volume as Class increases until class A4. After class A4 we see a decrease in Shuck from A4 to A5. Questions that arise: 1) What factors are causing the Shuck to peak and then decrease from class A4 to A5? ***

5(c) (3 points) Present four boxplots using *par(mfrow = c(2, 2)* or *grid.arrange()*. The first line should show VOLUME by RINGS for the infants and, separately, for the adult; factor levels "M" and "F," combined. The second line should show WHOLE by RINGS for the infants and, separately, for the adults. Since the data are sparse beyond 15 rings, limit the displays to less than 16 rings. One way to accomplish this is to generate a new data set using subset() to select RINGS < 16.  Use ylim = c(0, 1100) for VOLUME and ylim = c(0, 400) for WHOLE.  If you wish to reorder the displays for presentation purposes or use ggplot2 go ahead.

```{r Part_5c}
par(bg="gray98")
par(mfrow = c(2, 2))
Subset1 <- subset(mydata,mydata$RINGS < 16)
MF <- factor(Subset1$SEX,c("M","F"))

boxplot(Subset1$VOLUME~Subset1$RINGS,subset = Subset1$SEX == "I",ylim = c(0, 1100),col="dodgerblue",
        xlab="Rings",ylab="Volume",main="Infant Volume | Rings")

boxplot(Subset1$VOLUME~Subset1$RINGS,MF,ylim = c(0, 1100),col="brown",
        xlab="Rings",ylab="Volume",main="Adult Volume | Rings")

boxplot(Subset1$WHOLE~Subset1$RINGS,subset = Subset1$SEX == "I",ylim = c(0, 400),col="dodgerblue",
        xlab="Rings",ylab="Whole weight",main="Infant Whole weight | Rings")

boxplot(Subset1$WHOLE~Subset1$RINGS,MF,ylim = c(0, 400),col="brown",
        xlab="Rings",ylab="Whole weight",main="Adult Whole weight | Rings")
```

**Essay Question (2 points):  What do these displays suggest about abalone growth?  Also, compare the infant and adult displays.  What differences stand out?**

***Answer: These displays indicate that for Infants, as the Ring count increases so does the median volume (up to 11 Rings). In addition to the median increase, we also can see a greater dispersion of the quartiles. In particular, the interquartile range grows up to 12 before we see a reverse effect (decrease). When we look at the Infant weight, we also see that weight peaks in the 12 rings range. Here again we see a larger interquartile range at its peak before it recedes in size for Rings 13-15. Adults reflect s similar growth pattern; however, the sample ranges are much larger. In adult abalones we see a peak range of volume and Rings when the Ring sizes are between 11-12. When compared to Infants we can observe that adult abalones reflect greater consistency across Rings and volume size peak. Additionally, they also have a large range in volume and weight.***

-----

##### Section 6: (11 points) Conclusions from the Exploratory Data Analysis (EDA).

**Conclusions**


Essay Question 1) (5 points)	Based solely on these data, what are plausible statistical reasons that explain the failure of the original study? Consider to what extent physical measurements may be used for age prediction.

***Answer: The data contains outliers as evidenced by boxplots. It is plausible that one of the failures of the original study occurred because of abnormalities such as outliers or incorrect measurements (Random Error) were used to determine the age of an Abalone. This may have been a scenario where there was correlation but not causation. Additionally, there are other challenges that could have led to misclassified data such as ring count and sexing. Which brings to question, “what kind of processes were used to observe?” and “was the observation/sample representative of the population?”. It is important that the process used for data collection is consistent and as stated before, representative of the population. Considering the physical measurements and what may have been used for age prediction it would require a repeatable and reproducible  measurement process that is applied to all specimens for measurement. Failure to do this would result in inaccurate results. Also, the stratification of data to determine classes would be especially important. Failure to properly classify an Abalone can also lead to inaccurate results. ***

Essay Question 2) (3 points) Do not refer to the abalone data or study.  If you were presented with an overall histogram and summary statistics from a sample of some population or phenomenon and no other information, what questions might you ask before accepting them as representative of the sampled population or phenomenon?

***Answer: It is important to understand how the sample was obtained and if any consideration was given to the potential for bias. In this situation, it is important to have a sample that represents the population without influence or “bias”. For example, questions would include: 1) How was the sample determined? This would clarify if any consideration was given to bias. 2) Was probability sampling used? This would provide feedback on the representativeness of the sample obtained compared to the overall population. It would be important to know that appropriate steps such as Sample Random Sampling (SRS) are used to avoid any biases in the sample 3) How was the sampling size determined? This would identify what method was used to identify the sample size. If this was not used, then you could be shown data that has an insufficient representation of the population. 5) Hoe were outliers addressed? It is important to understand if outliers are removed. Failure to remove/address outliers can result in a skewed histogram and provide inaccurate results.***

Essay Question 3)  (3 points)	Do not refer to the abalone data or study.  What do you see as difficulties analyzing data derived from observational studies? Can causality be determined?  What might be learned from such studies?

***Answer: Observational studies are based on observations without manipulation. Therefore, data collection can be susceptible to researcher bias, lack of observational structure (i.e. process used), or their mis-interpretation of the sample. For example, missed observations due to a poor structured process can lead to the non-collection of critical data (or the collection of data that does not properly represent the observation/sample). Furthermore, incorrect data (or lack thereof) can lead to incorrect information (causality). However, as discussed during the semester, correlation does not mean causation. I think an important take away from such studies is understanding the criteria and underlying activities used in an observation. How the information was gathered and what steps were taken to observe the process are important factors when taking into consideration data from a study. In six sigma we use a process called Gage which examines the repeatability and reproducibility of a measurement. This is a critical factor in any process because it addresses the method that was taken to gather information/measurements and how it compares when the process is reproduced (checks for variation). Similarly, any information collected from observational studies can potentially have inconsistent practices and biases which can lead to causation. A lesson from such studies is that it is a best practice to understand the methods that were used in that study and be objective. ***
