---
title: "Data Analysis #2 Version 2 (75 points total)"
author: "Mark, Ryan"
output:
  html_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```
### Project Overview: ###

***This project will examine sample data collected from a large marine sea snail called Abalone. The information that we have retrieved is based on an observational study that was performed on Abalones. The objective of that study was to be able to predict the age of Abalones based on their physical measurements. At the conclusion of that study it was determined that it was unsuccessful at meeting its overall objective. The intent of this exploratory analysis is to determine plausible reasons why the original study was unsuccessful in predicting abalone age based on physical characteristics.***

##Data Analysis #2

```{r analysis_setup1, message = FALSE, warning = FALSE}

# Perform the following steps to start the assignment.
 
# 1) Load/attach the following packages via library():  flux, ggplot2, gridExtra, moments, rockchalk, car.
# NOTE:  packages must be installed via install.packages() before they can be loaded.

library(flux)
library(ggplot2)
library(gridExtra)
library(moments)
library(rockchalk) # base R code replaces requirement for this package
library(car)

# 2) Use the "mydata.csv" file from Assignment #1 or use the file posted on the course site.  Reading
# the files into R will require sep = "" or sep = " " to format data properly.  Use str() to check file
# structure.

mydata <- read.csv("mydata.csv", sep = ",")
# mydata <- read.csv(file.path("c:...", "mydata.csv"), sep = ",")
# mydata <- read.csv(file.path("c:/Rabalone/", "mydata.csv"), sep = ",")

str(mydata)

```

### Test Items starts from here - There are 10 sections - total of 75 points ##############

##### Section 1: (5 points)

(1)(a) Form a histogram and QQ plot using RATIO. Calculate skewness and kurtosis using 'rockchalk.' Be aware that with 'rockchalk', the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_1a}
mydata$VOLUME <- mydata$LENGTH * mydata$DIAM * mydata$HEIGHT
mydata$RATIO <- mydata$SHUCK/mydata$VOLUME

par(mfrow=c(1,2))
hist(mydata$RATIO,
     main="Histogram of Ratio",
     xlab="Ratio",
     col="yellow")

qqnorm(mydata$RATIO,
       main="QQ Plot",
       col="green3")

qqline(mydata$RATIO,
       col="red")

rockchalk::skewness(mydata$RATIO)
rockchalk::kurtosis(mydata$RATIO,excess=FALSE) # Used false to capture an equivalent to the moments pkg.
```

(1)(b) Tranform RATIO using *log10()* to create L_RATIO (Kabacoff Section 8.5.2, p. 199-200). Form a histogram and QQ plot using L_RATIO. Calculate the skewness and kurtosis. Create a boxplot of L_RATIO differentiated by CLASS.

```{r Part_1b}
mydata$L_Ratio <- log10(mydata$RATIO)

par(mfrow=c(1,3))
hist(mydata$L_Ratio,main="Histogram of L_Ratio",
     xlab="L_Ratio",
     col="green4")

qqnorm(mydata$L_Ratio,main="QQ Plot",
       col="blue")
qqline(mydata$L_Ratio,
       col="red")

boxplot(mydata$L_Ratio~mydata$CLASS,
        main="Boxplot of L_Ratio to Class",
        xlab="by Class",ylab="Frequency",
        col=c("red","yellow","green","blue","purple"))
```

(1)(c) Test the homogeneity of variance across classes using *bartlett.test()* (Kabacoff Section 9.2.2, p. 222). 

```{r Part_1c}
x <- bartlett.test(RATIO~CLASS,data=mydata)
y <- bartlett.test(L_Ratio~CLASS,data=mydata)

print(x)
print(y)
```

**Essay Question: Based on steps 1.a, 1.b and 1.c, which variable RATIO or L_RATIO exhibits better conformance to a normal distribution with homogeneous variances across age classes?  Why?** 

***Answer: The observations of each graph indicate that the L_Ratio exhibits a better conformance to a normal distribution than Ratio. This is further proved through the use of the Bartlett test which indicates that we fail to reject the null hypothesis (Variances are equal).***


##### Section 2 (10 points) ###############################

(2)(a) Perform an analysis of variance with *aov()* on L_RATIO using CLASS and SEX as the independent variables (Kabacoff chapter 9, p. 212-229). Assume equal variances. Perform two analyses. First, fit a model with the interaction term CLASS:SEX. Then, fit a model without CLASS:SEX. Use *summary()* to obtain the analysis of variance tables (Kabacoff chapter 9, p. 227).

```{r Part_2a}
summary(aov(L_Ratio~CLASS+SEX+CLASS:SEX,data=mydata))
summary(aov(L_Ratio~CLASS+SEX,data=mydata))
```

**Essay Question:  Compare the two analyses.  What does the non-significant interaction term suggest about the relationship between L_RATIO and the factors CLASS and SEX?**

***Answer: Based on the evaluation of the P-value we are able to determine that the interaction of Class:Sex is insignificant. This is indicated by a P-value of .867. When we evaluate the individual variables of Class and Sex we can conclude by their P-values the significance.***

(2)(b) For the model without CLASS:SEX (i.e. an interaction term), obtain multiple comparisons with the *TukeyHSD()* function. Interpret the results at the 95% confidence level (*TukeyHSD()* will adjust for unequal sample sizes). 

```{r Part_2b}
TukeyHSD(aov(L_Ratio~CLASS+SEX,data=mydata))
```

**Additional Essay Question:  first, interpret the trend in coefficients across age classes. What is this indicating about L_RATIO?  Second, do these results suggest male and female abalones can be combined into a single category labeled as 'adults?' If not, why not?**

***Answer: When evaluating classes, we can see based on the Tukey HSD P-value that there is a statistical significance between classes A2-A1. All other class comparisons of means reflect that there is not a statistically significant difference. When we apply this evaluation to Sex it can be observed that I-F and M_I have a P-value <.05 which results in rejecting the null hypothesis. Therefore, we can conclude that male and female can be combined into a group ("Adult").***


######  Section 3: (10 points) ##################

(3)(a1) We combine "M" and "F" into a new level, "ADULT". (While this could be accomplished using *combineLevels()* from the 'rockchalk' package, we use base R code because many students do not have access to the rockchalk package.) This necessitated defining a new variable, TYPE, in mydata which had two levels:  "I" and "ADULT". 

```{r Part_3a1}
# here we show how to define the new variable TYPE using only base R functions (no need for outside packages)
mydata$TYPE <- character(nrow(mydata))  # initialize the TYPE column as all blanks
for (i in seq(along = mydata$SEX)) {
  mydata$TYPE[i] <- 'I'
  if (mydata$SEX[i] == 'M' || mydata$SEX[i] == 'F') mydata$TYPE[i] <- 'ADULT'
}
mydata$TYPE <- factor(mydata$TYPE)
cat('\nCheck on definition of TYPE object (should be an integer): ', typeof(mydata$TYPE))
cat('\nmydata$TYPE is treated as a factor: ', is.factor(mydata$TYPE), '\n')
table(mydata$SEX, mydata$TYPE)
```
(3)(a2)  Present side-by-side histograms of VOLUME. One should display infant volumes and, the other, adult volumes. 

```{r Part_3a2}
par(mfrow=c(1,2))

hist(mydata$VOLUME[mydata$TYPE=='I'],
     main="Infant Volumes",
     xlab="Volume",
     col="aquamarine")

hist(mydata$VOLUME[mydata$TYPE=='ADULT'],
     main="Adult Volumes",
     xlab='Volume',
     col="purple")
```


**Essay Question: Compare the histograms.  How do the distributions differ? Are there going to be any difficulties separating infants from adults based on VOLUME?**

***Answer: When we compare the distributions of Infants to Adults it is observed that Infants reflect a right skewed distribution. The Adult distribution reflects a more proportional/balanced distribution. When evaluating the Volume, we can see that the Infants have a higher distribution up to 200 whereas the Adults reflect peak ranges from 300-600. Because of the separation in volume this should be conducive to separating the two.***

(3)(b) Create a scatterplot of SHUCK versus VOLUME and a scatterplot of their base ten logarithms, labeling the variables as L_SHUCK and L_VOLUME. Please be aware the variables, L_SHUCK and L_VOLUME, present the data as orders of magnitude (i.e. VOLUME = 100 = 10^2 becomes L_VOLUME = 2). Use color to differentiate CLASS in the plots. Repeat using color to differentiate by TYPE. 

```{r Part_3b}
library(dplyr)
library(ggplot2)

mydata <- mydata %>% mutate(L_SHUCK = log10(SHUCK), L_VOLUME = log10(VOLUME))

a <- ggplot(mydata, aes(VOLUME, SHUCK)) + 
  theme_classic()
b <- ggplot(mydata, aes(L_VOLUME, L_SHUCK)) + 
  theme_classic()

gridExtra::grid.arrange(
  a + geom_point(aes(fill = CLASS), 
                 shape = 21),
  b + geom_point(aes(fill = CLASS), 
                 shape = 21),
  a + geom_point(aes(fill = TYPE) , 
                 shape = 21),
  b + geom_point(aes(fill = TYPE) , 
                 shape = 21)
)
```

**Additional Essay Question:  Compare the two scatterplots. What effect(s) does log-transformation appear to have on the variability present in the plot?  What are the implications for linear regression analysis? Where do the various CLASS levels appear in the plots? Where do the levels of TYPE appear in the plots?**

***Answer: The log-transformation appears to have established better visibility of the clusters. The visual aspect of this graph allows us to see an improved variability between clusters whereas the Shuck vs. Volume contained a thicker range of clusters with the potential for overlap. Furthermore, we can see separation between classes and type with the use of log-transformation. This makes it easier to identify (visually) between types or class.***


######   Section 4: (5 points) ###################################

(4)(a1) Since abalone growth slows after class A3, infants in classes A4 and A5 are considered mature and candidates for harvest. Reclassify the infants in classes A4 and A5 as ADULTS. This reclassification could have been achieved using *combineLevels()*, but only on the abalones in classes A4 and A5. We will do this recoding of the TYPE variable using base R functions. We will use this recoded TYPE variable, in which the infants in A4 and A5 are reclassified as ADULTS, for the remainder of this data analysis assignment. 

```{r Part_4a1}
for (i in seq(along = mydata$TYPE)) {
  if (mydata$CLASS[i] == 'A4' || mydata$CLASS[i] == 'A5') mydata$TYPE[i] <- 'ADULT'
}
mydata$TYPE <- factor(mydata$TYPE)
cat('\nCheck on redefinition of TYPE object (should be an integer): ', typeof(mydata$TYPE))
cat('\nmydata$TYPE is treated as a factor: ', is.factor(mydata$TYPE), '\n')
cat('\nThree-way contingency table for SEX, CLASS, and TYPE:\n')
print(table(mydata$SEX, mydata$CLASS, mydata$TYPE))
```

(4)(a2) Regress L_SHUCK as the dependent variable on L_VOLUME, CLASS and TYPE (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2 and Black Section 14.2). Use the multiple regression model: L_SHUCK ~ L_VOLUME + CLASS + TYPE. Apply *summary()* to the model object to produce results.

```{r Part_4a2}
model <- lm(L_SHUCK~L_VOLUME+CLASS+TYPE,data = mydata)
summary(model)
```

**Essay Question:  Interpret the trend in CLASS levelcoefficient estimates? (Hint:  this question is not asking if the estimates are statistically significant. It is asking for an interpretation of the pattern in these coefficients, and how this pattern relates to the earlier displays).**

***Answer: The coefficient data indicates a negative increase in the estimates as the class increases. This would indicate that as the class number increases there is a negative increase for the L_Shuck.***

**Additional Essay Question:  Is TYPE an important predictor in this regression? (Hint:  This question is not asking if TYPE is statistically significant, but rather how it compares to the other independent variables in terms of its contribution to predictions of L_SHUCK for harvesting decisions.)  Explain your conclusion.**

***Answer: The type is not a strong indicator as evidenced by the coefficient value of L_Shuck and when compared to classes and volume. Therefore, this may not be a strong contribution for predicting _Shuck for harvesting decisions.***

-----

The next two analysis steps involve an analysis of the residuals resulting from the regression model in (4)(a) (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2).

-----

###### Section 5: (5 points) #################################

(5)(a) If "model" is the regression object, use model$residuals and construct a histogram and QQ plot. Compute the skewness and kurtosis. Be aware that with 'rockchalk,' the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_5a}
rockchalk::skewness(model$residuals)
rockchalk::kurtosis(model$residuals, excess = FALSE) # Used false to capture an equivalent to the moments pkg.

par(mfrow = c(1, 2))

hist(model$residuals,
     main = "Histogram of Model_Residuals",
     xlab = "Model_Residuals",
     col = "violet")

qqnorm(model$residuals,
       main = "QQ Plot",
       col="green4")

qqline(model$residuals,col="red")
```

(5)(b) Plot the residuals versus L_VOLUME, coloring the data points by CLASS and, a second time, coloring the data points by TYPE. Keep in mind the y-axis and x-axis may be disproportionate which will amplify the variability in the residuals. Present boxplots of the residuals differentiated by CLASS and TYPE (These four plots can be conveniently presented on one page using *par(mfrow..)* or *grid.arrange()*. Test the homogeneity of variance of the residuals across classes using *bartlett.test()* (Kabacoff Section 9.3.2, p. 222).  

```{r Part_5b}
library(ggplot2)

mydata$RESIDUALS <- model$residuals

gridExtra::grid.arrange(
  ggplot(mydata, aes(L_VOLUME, mydata$RESIDUALS)) + 
    theme_classic() + 
    geom_point(aes(fill = CLASS), shape = 21),
  ggplot(mydata, aes(L_VOLUME, mydata$RESIDUALS)) + 
    theme_classic() + 
    geom_point(aes(fill = TYPE),  shape = 21),
  ggplot(mydata, aes(CLASS, mydata$RESIDUALS))    + 
    theme_classic() + 
    geom_boxplot(aes(fill = CLASS)),
  ggplot(mydata, aes(TYPE, mydata$RESIDUALS))     + 
    theme_classic() + 
    geom_boxplot(aes(fill = TYPE))
)

bartlett.test(mydata$RESIDUALS ~ CLASS, data = mydata)
```

**Essay Question:  What is revealed by the displays and calculations in (5)(a) and (5)(b)? Does the model 'fit'?  Does this analysis indicate that L_VOLUME, and ultimately VOLUME, might be useful for harvesting decisions? Discuss.**  

***Answer: The histogram created in section 5a or residuals reflects a normal distribution. When we examine the boxplots in section 5b we can see the similar quartile ranges for residuals by class. When looking at the clusters of residuals to L_volume we can observe scattered variability with a higher concentration favoring the right side of the graph.***

-----

There is a tradeoff faced in managing abalone harvest. The infant population must be protected since it represents future harvests. On the other hand, the harvest should be designed to be efficient with a yield to justify the effort. This assignment will use VOLUME to form binary decision rules to guide harvesting. If VOLUME is below a "cutoff" (i.e. a specified volume), that individual will not be harvested. If above, it will be harvested. Different rules are possible.

The next steps in the assignment will require consideration of the proportions of infants and adults harvested at different cutoffs. For this, similar "for-loops" will be used to compute the harvest proportions. These loops must use the same values for the constants min.v and delta and use the same statement "for(k in 1:10000)."  Otherwise, the resulting infant and adult proportions cannot be directly compared and plotted as requested. Note the example code supplied below.

-----

#### Section 6: (5 points) ########################

(6)(a) A series of volumes covering the range from minimum to maximum abalone volume will be used in a "for loop" to determine how the harvest proportions change as the "cutoff" changes. Code for doing this is provided.

```{r Part_6a}

idxi <- mydata$TYPE == "I"
idxa <- mydata$TYPE == "ADULT"

max.v <- max(mydata$VOLUME)
min.v <- min(mydata$VOLUME)
delta <- (max.v - min.v)/10000
prop.infants <- numeric(10000)
prop.adults <- numeric(10000)
volume.value <- numeric(10000)

total.infants <- sum(idxi)  
total.adults <- sum(idxa)

for (k in 1:10000) { 
	value <- min.v + k*delta
	volume.value[k] <- value
	prop.infants[k] <- sum(mydata$VOLUME[idxi] <= value)/total.infants
	prop.adults[k] <-  sum(mydata$VOLUME[idxa] <= value)/total.adults
}

# prop.infants shows the impact of increasing the volume cutoff for
# harvesting. The following code shows how to "split" the population at
# a 50% harvest of infants.

n.infants <- sum(prop.infants <= 0.5)
split.infants <- min.v + (n.infants + 0.5)*delta  # This estimates the desired volume.
split.infants

n.adults <- sum(prop.adults <= 0.5)
split.adults <- min.v + (n.adults + 0.5)*delta
split.adults
```

(6)(b) Present a plot showing the infant proportions and the adult proportions versus volume.value. Compute the 50% "split" volume.value for each and show on the plot.   

```{r Part_6b}
library(ggplot2)

ggplot() + theme_bw() +
  geom_line(aes(volume.value, prop.infants), 
            color = "red") +
  geom_line(aes(volume.value, prop.adults), 
            color = "purple") + 
  geom_vline(xintercept = split.infants) +
  annotate("text", label = paste(round(split.infants, 3), "\nInfants(I)"), 
           x = split.infants + 
             80, y = 0.57) +
  geom_vline(xintercept = split.adults) +
  annotate("text", label = paste(round(split.adults, 3), "\nAdults(A)"), 
           x = split.adults + 
             110, y = 0.57) +
  geom_hline(yintercept = 0.5) +
  labs(x = "Volume", y = "Proportion", 
       title = "The Infant proportions and the Adult proportions vs. volume")
```

**Essay Question:  The two 50% "split" values serve a descriptive purpose illustrating the difference between the populations. What do these values suggest regarding possible cutoffs for harvesting?** 

***Answer: It indicates that each cutoff point is good for harvesting. We can tell based on the differences in proportion and volume of the cutoffs.***

-----

This part will address the determination of a volume.value corresponding to the observed maximum difference in harvest percentages of adults and infants. To calculate this result, the vectors of proportions from item (6) must be used. These proportions must be converted from "not harvested" to "harvested" proportions by using (1 - prop.infants) for infants, and (1 - prop.adults) for adults. The reason the proportion for infants drops sooner than adults is that infants are maturing and becoming adults with larger volumes.

-----

###### Section 7: (10 points)  #######################

(7)(a) Evaluate a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value. Compare to the 50% "split" points determined in (6)(a). There is considerable variability present in the peak area of this plot. The observed "peak" difference may not be the best representation of the data. One solution is to smooth the data to determine a more representative estimate of the maximum difference.

```{r Part_7a}
library(ggplot2)

difference <- ((1 - prop.adults) - (1 - prop.infants))

ggplot() + theme_classic() +
  geom_line(aes(volume.value, difference),color = "red") +
  geom_vline(xintercept = split.infants) +
  annotate("text", label = paste('*Infant vol. =', round(split.infants, 3)), 
           x = split.infants + 130, y = 0.3) +
  geom_vline(xintercept = split.adults) +
  annotate("text", label = paste('*Adult vol. =', round(split.adults, 3)), 
           x = split.adults + 140, y = 0.05) +
  labs(x = "Volume", y = "Difference", 
       title = "Difference in Proportions")
```

(7)(b) Since curve smoothing is not studied in this course, code is supplied below. Execute the following code to create a smoothed curve to append to the plot in (a). The procedure is to individually smooth (1-prop.adults) and (1-prop.infants) before determining an estimate of the maximum difference. 

```{r Part_7b}
y.loess.a <- loess(1 - prop.adults ~ volume.value, span = 0.25,
	family = c("symmetric"))
y.loess.i <- loess(1 - prop.infants ~ volume.value, span = 0.25,
	family = c("symmetric"))
smooth.difference <- predict(y.loess.a) - predict(y.loess.i)
```

(7)(c) Present a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value with the variable smooth.difference superimposed. Determine the volume.value corresponding to the maximum smoothed difference (Hint:  use *which.max()*). Show the estimated peak location corresponding to the cutoff determined.

```{r Part_7c}
library(ggplot2)

cutoff1 <- volume.value[which.max(smooth.difference)]
paste('cutoff1 =',round(cutoff1,3))

ggplot() + theme_classic() +
  geom_line(aes(volume.value, difference), color = "red") +
  geom_line(aes(volume.value, smooth.difference), color = "blue") +
  geom_vline(xintercept = volume.value[cutoff1]) +
  annotate("text", label = paste('Volume =', round(cutoff1, 3)), 
           x = cutoff1 + 20, y = 0.35) +
  labs(x = "VOLUME", y = "Difference in Proportions", 
       title = "Difference in Proportions")
```

(7)(d) What separate harvest proportions for infants and adults would result if this cutoff is used? Show the separate harvest proportions (NOTE:  the adult harvest proportion is the "true positive rate" and the infant harvest proportion is the "false positive rate").

Code for calculating the adult harvest proportion is provided.

```{r Part_7d}
(1 - prop.adults)[which.max(smooth.difference)]  # [1] 0.7416332
round((1 - prop.infants)[which(volume.value == cutoff1)],3)
```

-----

There are alternative ways to determine cutoffs. Two such cutoffs are described below.

-----

######  Section 8: (10 points)  ###################

(8)(a) Harvesting of infants in CLASS "A1" must be minimized. The smallest volume.value cutoff that produces a zero harvest of infants from CLASS "A1" may be used as a baseline for comparison with larger cutoffs. Any smaller cutoff would result in harvesting infants from CLASS "A1."  

Compute this cutoff, and the proportions of infants and adults with VOLUME exceeding this cutoff. Code for determining this cutoff is provided. Show these proportions.

```{r Part_8a}
library(dplyr)

volume.value[volume.value > max(mydata[mydata$CLASS == "A1" & mydata$TYPE == "I", "VOLUME"])][1] # [1] 206.786

a1i <- with(mydata, max(VOLUME[CLASS == 'A1' & TYPE == 'I']))
round(a1i,3)

cutoff2 <- first(volume.value[volume.value > a1i])
paste('cutoff2 =',round(cutoff2,3))

round((1 - prop.adults)[which(volume.value == cutoff2)],3)
round((1 - prop.infants)[which(volume.value == cutoff2)],3)
```

(8)(b) Another cutoff is one for which the proportion of adults not harvested equals the proportion of infants harvested. This cutoff would equate these rates; effectively, our two errors:  'missed' adults and wrongly-harvested infants. This leaves for discussion which is the greater loss:  a larger proportion of adults not harvested or infants harvested?  This cutoff is 237.7383. Calculate the separate harvest proportions for infants and adults using this cutoff. Show these proportions.  Code for determining this cutoff is provided.  

```{r Part_8b}
volume.value[which.min(abs(prop.adults - (1-prop.infants)))] # [1] 237.6391

cutoff3 <- volume.value[which.min(abs(prop.adults + prop.infants - 1))]
paste('cutoff3 =',round(cutoff3,3))

round((1 - prop.adults)[which(volume.value == cutoff3)],3)
round((1 - prop.infants)[which(volume.value == cutoff3)],3)
```


##### Section 9: (5 points) ###########

(9)(a) Construct an ROC curve by plotting (1 - prop.adults) versus (1 - prop.infants). Each point which appears corresponds to a particular volume.value. Show the location of the cutoffs determined in (7) and (8) on this plot and label each. 

```{r Part_9}
library(ggplot2)

allcutoffs <- which(volume.value %in% c(cutoff1, cutoff2, cutoff3))

ggplot(mapping = aes((1 - prop.infants), (1 - prop.adults))) + 
  geom_line(color = 'blue', size = 1) + theme_classic() +
  geom_abline(intercept = 0, slope = 1, color = 'red', linetype = 1) +
  geom_point(mapping = aes((1 - prop.infants[allcutoffs]), (1 - prop.adults[allcutoffs])),
             shape = 19, size = 3,color="red") +
  annotate("text", label = paste('equal error (vol) =', round(cutoff3, 3)), 
           x = 0.16, y = 0.87, size = 4) +
  annotate("text", label = paste('Zero harvest A1 infants (vol) =', round(cutoff2, 3)), 
           x = 0.45, y = 0.77, size = 4) +
  annotate("text", label = paste('Max Diff. (vol) =', round(cutoff1, 3)), 
           x = 0.30, y = 0.70, size = 4) +
  labs(title = "ROC curve ",
       x = "Infant proportion", y = "Adult proportion") + 
  theme(axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5))
```

(9)(b) Numerically integrate the area under the ROC curve and report your result. This is most easily done with the *auc()* function from the "flux" package.   Areas-under-curve, or AUCs, greater than 0.8 are taken to indicate good discrimination potential. 

```{r Part_9b}
round(flux::auc(x = (1 - prop.infants), y = (1 - prop.adults)),3)
```


##### Section 10: (10 points) ###################

(10)(a) Prepare a table showing each cutoff along with the following:
 	1) true positive rate (1-prop.adults,
 	2) false positive rate (1-prop.infants),
 	3) harvest proportion of the total population
 	
```{r Part_10} 	
True.Pos.Rate <- (1 - prop.adults)[allcutoffs]
round(True.Pos.Rate,3)

False.Pos.Rate <- (1 - prop.infants)[allcutoffs]
round(False.Pos.Rate,3)

Prop.of.ttl.pop <- (True.Pos.Rate * total.adults + False.Pos.Rate * total.infants) / (total.adults + total.infants)
round(Prop.of.ttl.pop,3)

tribble(
  ~strategy, ~volume, ~True.Pos.Rate, ~False.Pos.Rate, ~Prop.of.ttl.pop,
  "Max diff."  , cutoff1, True.Pos.Rate[3], False.Pos.Rate[3], Prop.of.ttl.pop[3],
  "Zero Harv. A1 infants" , cutoff2, True.Pos.Rate[1], False.Pos.Rate[1], Prop.of.ttl.pop[1],
  "equal Harv.", cutoff3, True.Pos.Rate[2], False.Pos.Rate[2], Prop.of.ttl.pop[2]
)
```
 	
**Essay Question: Based on the ROC curve, it is evident a wide range of possible "cutoffs" exist. Compare and discuss the three cutoffs determined in this assignment.**   

***Answer: Max difference reflects the lowest false positive rate and the highest cutoff. The Zero Harvest A1 infants reflects the lowest volume cutoff, highest True-positive-rate, and the highest false-positive-rate. The equal Harvest reflects the middle cutoff, middle True-positive-rate, and middle false-positive-rate.***


**Final Essay Question:  Assume you are expected to make a presentation of your analysis to the investigators How would you do so?  Consider the following in your answer:**

1. Would you make a specific recommendation or outline various choices and tradeoffs?
2. What qualifications or limitations would you present regarding your analysis?
3. If it is necessary to proceed based on the current analysis, what suggestions would you have for implementation of a cutoff?  4)  What suggestions would you have for planning future abalone studies of this type? 

***Answer: My approach would be to address the process used and how information was root caused. I would use this approach because I would be able to explain how I arrived at my findings and what conclusions or recommendations would be made as a result of those findings. For example, this process could start with the initial analysis of data by sex, class, ratio, etc. I would address limitations by discussing what analyses were helpful and what did not generate information that was needed. If it is necessary to proceed, I believe equal harvesting would offer a median approach that does not over-harvest nor susceptible to high and low True/false rates when compared to the other cutoffs. This is a neutral approach that avoids the potential pitfalls  seen at higher rates with the other cutoffs. For planning future abalone studies, I believe focusing on an optimal measurement plan would benefit this process greatly. This approach would focus on the value of measurements by examining the repeatability and reproducibility.  The goal of this plan would be to optimize how we measure the age of Abalones so that it is sustainable and accurate.***